Synthesis of Rules for Network Intrusion Detectors

### 2019-10-24 paper/paper.tex
pass on recon
pass on pseudo-code

\begin{document}
%
\title{Synthesis of Rules for Network Intrusion Detectors}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Lucas Alcantara\MyComment{\inst{1}}\MyComment{\orcidID{0000-1111-2222-3333}} \and
Guilherme Padilha\MyComment{\inst{1}}\MyComment{\orcidID{1111-2222-3333-4444}} \and
Marcelo d'Amorim\MyComment{\orcidID{0000-0002-1323-8769}}}
%
\authorrunning{Alcantara et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Federal University of Pernambuco (UFPE), Recife, Brazil\\
\email{\{lama,ghps,damorim\}@cin.ufpe.br}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Network Intrusion Detection Systems (\nids{}) are a popular mechanism
used by system administrators to defend against network attacks. These
systems monitor the network traffic and flag suspicious network
behavior. Signature-based \nids\ do that by checking the network
traffic against a pre-defined set of rules, which can become obsolete
as attackers learn new strategies to circumvent existing defenses.
This paper proposes \tname{}\footnote{Abbreviation of \textbf{Sy}nthesis of
  Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a technique that automatically
synthesizes \nids\ rules from positive and negative examples (\ie{},
malicious and benign traffic). \tname{} formulates synthesis as an
optimization problem whose candidate solutions are rules that
maximizes the captured the positive traffic and minimizes the captured
negative traffic. \tname{} bootstraps the search with candidate
solutions that use data from the payload of the malicious message and
produces optimal rule candidates on output. We evaluated \tname{} on a
diverse set of attacks. Results indicate that \Fix{...}

@@

\tname{} is a technique that uses positive and negative examples to
synthesize signature-based \nids rules, such as those from Snort and
Suricata. The goal of \tname\ is to create a rule that \emph{only}
captures the positive traffic, \ie{}, the malicious sequence of
messages.

@@

\tname{} leverages public datasets\Fix{cite cite} of negative traffic
to guide a search for more general rules, hence addressing the
overffiting issue. The goal of this step is to create rules that
captures traffic similar to the input malicious traffic. To achieve
that goal, \tname{} \emph{minimizes} the rules produced in the
previous step, preserving the invariant of \emph{only} capturing the
positive malicious traffic. A rule is minimal if discarding any option
results in capturing negative (\ie{}, benign) traffic.  The
minimization procedure we propose in this step builds on the
observation that all options from the rule produced in the previous
step are satisifed by the positive traffic. Consequently, any subset
of these options also captures the positive traffic. The procedure
iteratively discards options from the candidate rule until it detects
that the rule is minimal, \ie{}, it finds that no other options can be
removed. The procedure uses negative traffic to guide the search.

\begin{algorithm}[t!]
\textbf{INPUT:} overconstrained rule $or$ \\
\textbf{OUTPUT:} Set of minimal rules $R$\\
\vspace{-3ex}
%\noindent\begin{minipage}{\textwidth}
\captionof{algorithm}{Minimization algorithm}\label{algo1}
\begin{algorithmic}[1]
%  \Procedure{MinRule}{$or$}\Comment{The g.c.d. of a and b}
    \State $r\gets a\bmod b$
    \While{$r\not=0$}\Comment{We have the answer if r is 0}
      \State $a\gets b$
      \State $b\gets r$
      \State $r\gets a\bmod b$
    \EndWhile
    \State \textbf{return} $b$\Comment{The gcd is b}
%  \EndProcedure
\end{algorithmic}
%\end{minipage}%
\end{algorithm}

%% Public databases of negative traffic with
%% \Fix{thousands} of messages exist  and can be used to
%% increase the accuracy of the minimization.

\subsection{Limitations}

>>>>

\begin{document}
%
\title{Synthesis of Rules for Network Intrusion Detectors}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Lucas Alcantara\MyComment{\inst{1}}\MyComment{\orcidID{0000-1111-2222-3333}} \and
Guilherme Padilha\MyComment{\inst{1}}\MyComment{\orcidID{1111-2222-3333-4444}} \and
Marcelo d'Amorim\MyComment{\orcidID{0000-0002-1323-8769}}}
%
\authorrunning{Alcantara et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Federal University of Pernambuco (UFPE), Recife, Brazil\\
\email{\{lama,ghps,damorim\}@cin.ufpe.br}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Network Intrusion Detection Systems (\nids{}) are a popular mechanism
used by system administrators to defend against network attacks. These
systems monitor the network traffic and flag suspicious network
behavior. Signature-based \nids\ do that by checking the network
traffic against a pre-defined set of rules, which can become obsolete
as attackers learn new strategies to circumvent existing defenses.
This paper proposes \tname{}\footnote{Abbreviation of \textbf{Sy}nthesis of
  Su\textbf{ri}cata R\textbf{u}le\textbf{s}.}, a technique that automatically
synthesizes \nids\ rules from positive and negative examples (\ie{},
malicious and benign traffic). \tname{} formulates synthesis as an
optimization problem whose candidate solutions are rules that
maximizes the captured the positive traffic and minimizes the captured
negative traffic. \tname{} bootstraps the search with candidate
solutions that use data from the payload of the malicious message and
produces optimal rule candidates on output. We evaluated \tname{} on a
diverse set of attacks. Results indica te that \Fix{...}

@@

\tname{} is a technique that uses positive and negative examples to
synthesize signature-based \nids\ rules, such as those from Snort and
Suricata. The goal of \tname\ is to create a rule that \emph{only}
captures the positive traffic, \ie{}, the malicious sequence of
messages.

@@

\tname{} leverages public datasets\cite{tcpreplay} of negative traffic
to guide the search for more general rules, hence addressing the
overffiting issue. The goal of this step is to create rules that
captures traffic similar to the input malicious traffic. To achieve
that goal, \tname{} \emph{minimizes} the rules produced in the
previous step, preserving the invariant of \emph{only} capturing the
positive malicious traffic. A rule is minimal if discarding any option
results in capturing negative (\ie{}, benign) traffic.  The
minimization procedure we propose in this step builds on the
observation that all options from the rule produced in the previous
step are satisifed by the positive traffic. Consequently, any subset
of these options also captures the positive traffic. The procedure
iteratively discards options from the candidate rule until it detects
that the rule is minimal, \ie{}, it finds that no other options can be
removed. The procedure uses negative traffic to guide the search.

\algdef{SE}[DOWHILE]{Do}{doWhile}{\algorithmicdo}[1]{\algorithmicwhile\ #1}%
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\algnewcommand\And{\textbf{and} }
\begin{algorithm}[t!]
\textbf{INPUT:} overconstrained rule $or$ \\
\textbf{OUTPUT:} Set of minimal rules $\mathcal R$\\
\vspace{-3ex}
\captionof{algorithm}{Minimization algorithm}\label{algo1}
\begin{algorithmic}[1]
  \State $\mathcal R \gets \{\mathit{or}\}$\Comment{initial population
    of candidate solutions contains only the input rule.}\label{line-init}
  \Do\label{start-do-while}
      \State $\mathcal R' \gets \emptyset$\Comment{starting new
        iteration. new generation should be empty.}\label{init-rprime}
      \ForEach {$ro \in \mathcal R $}\Comment{analyze each candidate individual/rule separately.}\label{start-foreach}
           \For {$i \gets 1$ to $N$}\Comment{mutate each rule $N$ times.}
               \State $r \gets copy(ro)$      
               \State $j\gets \mathit{rand}(0,\mathit{len}(r))$
               \State $r[j]\gets 0$\Comment{discard option $j$ of rule
                 $r$, acopy of the original rule $\mathit{ro}$}
               \If{$\mathit{isnew}(r)$ \And $\mathit{isfit}(r)$}
                   \State $\mathcal R' \cup\gets \{r\}$\Comment{$r$
                     can lead to a plausible candidate solution.}\label{line-progress}
               \EndIf
           \EndFor
       \EndFor\label{end-foreach}
       \State $\mathcal R \gets R'$\Comment{update new generation}
  \doWhile{$\mathcal R'\not= \{\}$}\Comment{repeat until reaching a fixpoint}\label{end-do-while}
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{algo1} shows the pseudocode of the \tname{}
minimization algorithm. The algorithm takes as input the
overconstrained rule produced by the reverse engineering step and
produces a set $\mathcal R$ of rules that are minimal and only
captures the positive traffic. Line~\ref{line-init} initializes the
set $\mathcal{R}$ of candidate solutions with the input rule $or$. We
encode a rule as a bitvector where each index in the vector represents
one option; the value assigned at a given position indicates whether
or not the option is present in the rule. Each iteration of the outer
loop (lines~\ref{start-do-while}-\ref{end-do-while}) represents one
generation of the population of individuals produced by the
search. The algorithm terminates when no rule in the current
generation of individuals could be further reduced without violating
the property that a candidate solution should only capture positive
traffic. This means that execution has not reached
line~\ref{line-progress} as to augment the set $\mathcal{R'}$ and that
set will be empty at the end of one iteration of the outer
loop. Line~\ref{init-rprime} initializes the new generation of
candidate rules. The loop from
lines~\ref{start-foreach}--\ref{end-foreach} iterates through each
candidate rule in $\mathcal R$. Note that these rules have been found
in the previous generation. The algorithm mutates each of these rules
for $N$ times, discarding one option of the rule each time. The call
to $\mathit{isNew}$ checks if $r$ has not been previously generated
through another path whereas the call to $\mathit{isfit}$ checks if
the rule only captures positive traffic. Function $\mathit{copy}$
creates a copy of the bitvector of a rule, function $\mathit{len}$
returns the length of the vector, and function $\mathit{rand}$
randomly chooses one index of the vector.

%% Public databases of negative traffic with
%% \Fix{thousands} of messages exist  and can be used to
%% increase the accuracy of the minimization.
%\subsection{Limitations}
%\Fix{...}
